Solution-aware vs global ReLU selection: \\ partial MILP strikes back for DNN verification

### 2025-01-31 CAV 2025 TODO/introduction.tex
update and mark a problem

	\item On the first question, we adapted from BaB-SR \cite{BaB} and FSB \cite{FSB}, which choose branching nodes for BaB, global scoring (GS) functions to choose ReLU for pMILP. 
	These revealed to be much more accurate than the simple heuristic in 
	\cite{DivideAndSlide}. We uncover however one case (out of 4) where the exact function that GS tries to approximate depends heavily upon the mode of the ReLU function, giving very different result for each mode. This implies that these strategies may not be that accurate, because they do not know the mode to choose from. We then designed a novel solution-aware scoring (SAS), which uses the solution of a unique LP call, that provides the mode to consider. Theoretically, we show that SAS is always an over-approximation of the real improvement (Proposition \ref{prop2}), which implies that a small SAS value implies that the ReLU is unnecessary. Experimentally, we further show that SAS is very close to the actual improvement, closer than GS, and that overall, the accuracy from SAS is significantly better than GS.  Compared with the heuristic in \cite{DivideAndSlide}, SAS is much more efficient ($\approx 4$ times less integer variables for same accuracy (Table \ref{tab:example1})).We however explain why focusing on a solution may not be appropriate to choose branching order within this selection, and that GS would be better suited, which we verify experimentally.

>>>>

	\item On the first question, we adapted from BaB-SR \cite{BaB} and FSB \cite{FSB}, which choose branching nodes for BaB, global scoring (GS) functions to choose ReLU for pMILP. 
	These revealed to be much more accurate than the simple heuristic in 
	\cite{DivideAndSlide}. We uncover however one case (out of 4) where the exact function that GS tries to approximate depends heavily upon the mode of the ReLU function, giving very different result for each mode. This implies that these strategies may not be that accurate, because they do not know the mode to choose from. We then designed a novel solution-aware scoring (SAS), which uses the solution of a unique LP call, that provides the mode to consider. Theoretically, we show that SAS is always an over-approximation of the real improvement (Proposition \ref{prop2}), which implies that a small SAS value implies that the ReLU is unnecessary. Experimentally, we further show that SAS is very close to the actual improvement, closer than GS, and that overall, the accuracy from SAS is significantly better than GS.  Compared with the heuristic in \cite{DivideAndSlide}, {\color{red} SAS is much more efficient ($\approx 4$ times less integer variables for same accuracy (Table \ref{tab:example1}))}.We however explain why focusing on a solution may not be appropriate to choose branching order within this selection, and that GS would be better suited, which we verify experimentally.

### 2025-01-31 CAV 2025 TODO/proofsb.tex
update
update and mark a problem

Then, we explore the usefulness of computing accurately each layer inductively.
For that, we keep the setting of Figure \ref{fig3} / Table \ref{table14}, but computing the previous layer with LP rather than with full MILP.

@@

\section{SR Heuristic for choosing nodes}

Here we provide the formula of BaB-SR from paper \cite{BaB}.

\begin{align*}
	\nu_{n+1} &= -1&\\
	\hat{\nu}_k &= W^T_kv_{k+1}, k = n,\cdots,1 &\\
	\nu_{k,j} &= \begin{cases}
		0  &\text{ if }u_{k[j]}<0 \hspace*{2ex}(j\in\mathcal{I}^-_k)\\
		\hat{\nu}_{k[j]} &\text{ if }l_{k[j]}>0 \hspace*{2.5ex}(j\in\mathcal{I}^-_k)\\
		\frac{u_{k[j]}}{u_{k[j]}-l_{k[j]}}[\hat{\nu}_{k[j]}]_+-\frac{u_{k[j]}}{u_{k[j]}-l_{k[j]}}[\hat{\nu}_{k[j]}]_-  &\text{ otherwise } \hspace*{3ex}(j\in\mathcal{I}_k)\\ 
	\end{cases}\\
	&\hspace*{22ex}\text{for } k = n,\cdots,2&
\end{align*}and\begin{align*}
	s_{i[j]} = |\max(v_{i[j]}b_{i-1[j]},(v_{i[j]}-1)b_{i-1[j]})-\frac{u_{k[j]}}{u_{k[j]}-l_{k[j]}}[\hat{\nu}_{k[j]}]_+|
\end{align*}

\begin{figure}[b!]
	\hspace*{-0.8cm}
	\includegraphics[scale=0.6]{Plot for table 3}.
	\caption{Plot for table 3}
	\label{fig_table3}
\end{figure}

\begin{table}[h!]	
	\centering
	\begin{tabular}{|c||c|c|c|c|}
		\hline
		&\multicolumn{3}{c|}{$X \subseteq$ Layers 1\&2, max $=116$}\\\cline{2-5}
		\text{$|X|$} & \text{Random} &BSR &FSB &\bf Utility\\ \hline
		0  (LP) & 1.761  &  1.761& 1.761&1.761\\ \hline \hline
		5  & 1.729& 1.7138 & 1.7182 &\bf 1.532 \\ \hline
		10  & 1.696 & 1.6775&  1.6803&\bf  1.371\\ \hline
		15  &   1.653& 1.6392& 1.6475&\bf  1.247\\ \hline
		20  &   1.619 &1.5938 & 1.6174&\bf  1.145\\ \hline
		25  &    1.586 &1.5534 &1.5788 &\bf 1.061\\ \hline
		30  &  1.546 & 1.5166& 1.5413&\bf  0.989 \\ \hline
		35  &   1.502 & 1.4788 &1.4972 &\bf 0.934 \\ \hline
		40  &  1.469 & 1.4395& 1.4514&\bf 0.921 \\ \hline \hline
		max &  0.895 & 0.895 &0.895 &0.895   \\ \hline
	\end{tabular}
	\caption{BSR and FSB}
	\label{BSRandFSB}
	%\vspace{-0.6cm}
\end{table}

\subsection*{Formula for ResNet}

Consider a ResNet, like ResNet4b, we list the formula for an Add layer.

Suppose we are dealing with an Add layer $l$, target node $z$, with previous layer $l_1$ and shortcut connection previous layer $l_2$, starting from a layer $l_0$, that is:
\begin{align*}
	l &= l_1+l_2\\
	l_2 &= Conv_{W_2}(\ReLU(l_0))\\
	l_1' &= Conv_{W'_1}(\ReLU(l_0))\\
	l_1 &= Conv_{W_1}(\ReLU(l'_{1}))
\end{align*}The paths are $$l_0\rightarrow \ReLU(l_0)\rightarrow l_2\rightarrow l,$$ and $$l_0\rightarrow \ReLU(l_0)\rightarrow l'_1\rightarrow \ReLU(l_1')\rightarrow l_1\rightarrow l.$$

So, for nodes in layer $l_1'$, we can use the same formula as previous section, i.e., the formula at the end of page 6: for a node $b$ in $l_1'$, suppose $z=x+y$ for $x$ in layer $l_1$ and $y$ in $l_2$, 
\begin{align*}
\Utility\_\max\nolimits^x(b) &= W_{bx} \times (\sol(\hat{b})- \ReLU(\sol(b)))\\
\Utility\_\max\nolimits^z(b) &= \Utility\_\max\nolimits^x(b)
\end{align*}

For nodes in layer $l_0$, the simple idea is to consider both paths, and sum up the utilities of two paths. For a node $a$ in $l_0$, suppose $z=x+y$ for $x$ in layer $l_1$ and $y$ in $l_2$ ,
\begin{align*}
		\Delta(\hat{a}) &= \ReLU(\sol(a))-\sol(\hat{a})\\
		\Utility_\_\max\nolimits^y(a) &= -W_{ay} \Delta(\hat{a})\\
	\forall b \in \ell, \Delta(b) &= W_{ab}\Delta(\hat{a})\\
	\forall b \in \ell, \Delta(\hat{b}) &=
	\begin{cases}
		\frac{\UB(b)}{\UB(b)-\LB(b)}\Delta(b),  &\text{if }W_{bx} > 0\\
		\max(\Delta(b),-\sol(b)),  &\text{if }  W_{bx} < 0 \text{ and } \sol(b)\geq0\\
		\max(\Delta(b)+\sol(b),0),  &\text{if }  W_{bx} < 0 \text{ and } \sol(b)<0		 
	\end{cases}\\
	\Utility\_\max\nolimits^x(a) &= -\sum_{b \in \ell} W_{bx} \Delta(\hat{b})\\
	\Utility\_\max\nolimits^z(a) &= \Utility\_\max\nolimits^x(a)+\Utility\_\max\nolimits^y(a)
\end{align*}

\section{Formula for Utility}

Recall that for nodes $a,b$, we use 
$\hat{a}, \hat{b}$ to denote variable after ReLU functions, and
$\Delta(a),\Delta(\hat{a}),\Delta(b),\Delta(\hat{b})$ to denote the changes of those variables. $a$ means the source node, and $b$ means nodes one layer after $a$, and $z$ is 2 layers after $a$ and one after $b$.

The utility function for a node $a$ wrt neuron $z$ is defined inductively as follows:

\begin{align*}
	\Delta(\hat{a}) &= \ReLU(\sol(a))-\sol(\hat{a})\\
	\Delta(b) &= W_{ab}\Delta(\hat{a})\\
	\Delta(\hat{b}) &=
	\begin{cases}
		\frac{\UB(b)}{\UB(b)-\LB(b)}\Delta(b),  &\text{if }W_{bz} > 0\\
		\max(\Delta(b),-\sol(b)),  &\text{if }  W_{bz} < 0 \text{ and } \sol(b)\geq0\\
		\max(\Delta(b)+\sol(b),0),  &\text{if }  W_{bz} < 0 \text{ and } \sol(b)<0		 
	\end{cases}\\
	\Utility\_\max\nolimits^z(a) &= -\sum_{b \in \ell} W_{bz} \Delta(\hat{b})
\end{align*}

}

>>>>

Then, we explore the usefulness of computing accurately each layer inductively.
For that, we keep the setting of  Table \ref{table14}, but computing the previous layer with LP rather than with full MILP.

@@

%
%\section{SR Heuristic for choosing nodes}
%
%Here we provide the formula of BaB-SR from paper \cite{BaB}.
%
%\begin{align*}
%	\nu_{n+1} &= -1&\\
%	\hat{\nu}_k &= W^T_kv_{k+1}, k = n,\cdots,1 &\\
%	\nu_{k,j} &= \begin{cases}
%		0  &\text{ if }u_{k[j]}<0 \hspace*{2ex}(j\in\mathcal{I}^-_k)\\
%		\hat{\nu}_{k[j]} &\text{ if }l_{k[j]}>0 \hspace*{2.5ex}(j\in\mathcal{I}^-_k)\\
%		\frac{u_{k[j]}}{u_{k[j]}-l_{k[j]}}[\hat{\nu}_{k[j]}]_+-\frac{u_{k[j]}}{u_{k[j]}-l_{k[j]}}[\hat{\nu}_{k[j]}]_-  &\text{ otherwise } \hspace*{3ex}(j\in\mathcal{I}_k)\\ 
%	\end{cases}\\
%	&\hspace*{22ex}\text{for } k = n,\cdots,2&
%\end{align*}and\begin{align*}
%	s_{i[j]} = |\max(v_{i[j]}b_{i-1[j]},(v_{i[j]}-1)b_{i-1[j]})-\frac{u_{k[j]}}{u_{k[j]}-l_{k[j]}}[\hat{\nu}_{k[j]}]_+|
%\end{align*}
%
%
%\begin{figure}[b!]
%	\hspace*{-0.8cm}
%	\includegraphics[scale=0.6]{Plot for table 3}.
%	\caption{Plot for table 3}
%	\label{fig_table3}
%\end{figure}
%
%
%\begin{table}[h!]	
%	\centering
%	\begin{tabular}{|c||c|c|c|c|}
%		\hline
%		&\multicolumn{3}{c|}{$X \subseteq$ Layers 1\&2, max $=116$}\\\cline{2-5}
%		\text{$|X|$} & \text{Random} &BSR &FSB &\bf Utility\\ \hline
%		0  (LP) & 1.761  &  1.761& 1.761&1.761\\ \hline \hline
%		5  & 1.729& 1.7138 & 1.7182 &\bf 1.532 \\ \hline
%		10  & 1.696 & 1.6775&  1.6803&\bf  1.371\\ \hline
%		15  &   1.653& 1.6392& 1.6475&\bf  1.247\\ \hline
%		20  &   1.619 &1.5938 & 1.6174&\bf  1.145\\ \hline
%		25  &    1.586 &1.5534 &1.5788 &\bf 1.061\\ \hline
%		30  &  1.546 & 1.5166& 1.5413&\bf  0.989 \\ \hline
%		35  &   1.502 & 1.4788 &1.4972 &\bf 0.934 \\ \hline
%		40  &  1.469 & 1.4395& 1.4514&\bf 0.921 \\ \hline \hline
%		max &  0.895 & 0.895 &0.895 &0.895   \\ \hline
%	\end{tabular}
%	\caption{BSR and FSB}
%	\label{BSRandFSB}
%	%\vspace{-0.6cm}
%\end{table}
%
%
%
%\end{document}
%
%
%
%
%\subsection*{Formula for ResNet}
%
%Consider a ResNet, like ResNet4b, we list the formula for an Add layer.
%
%Suppose we are dealing with an Add layer $l$, target node $z$, with previous layer $l_1$ and shortcut connection previous layer $l_2$, starting from a layer $l_0$, that is:
%\begin{align*}
%	l &= l_1+l_2\\
%	l_2 &= Conv_{W_2}(\ReLU(l_0))\\
%	l_1' &= Conv_{W'_1}(\ReLU(l_0))\\
%	l_1 &= Conv_{W_1}(\ReLU(l'_{1}))
%\end{align*}The paths are $$l_0\rightarrow \ReLU(l_0)\rightarrow l_2\rightarrow l,$$ and $$l_0\rightarrow \ReLU(l_0)\rightarrow l'_1\rightarrow \ReLU(l_1')\rightarrow l_1\rightarrow l.$$
%
%So, for nodes in layer $l_1'$, we can use the same formula as previous section, i.e., the formula at the end of page 6: for a node $b$ in $l_1'$, suppose $z=x+y$ for $x$ in layer $l_1$ and $y$ in $l_2$, 
%\begin{align*}
%\Utility\_\max\nolimits^x(b) &= W_{bx} \times (\sol(\hat{b})- \ReLU(\sol(b)))\\
%\Utility\_\max\nolimits^z(b) &= \Utility\_\max\nolimits^x(b)
%\end{align*}
%
%For nodes in layer $l_0$, the simple idea is to consider both paths, and sum up the utilities of two paths. For a node $a$ in $l_0$, suppose $z=x+y$ for $x$ in layer $l_1$ and $y$ in $l_2$ ,
%\begin{align*}
%		\Delta(\hat{a}) &= \ReLU(\sol(a))-\sol(\hat{a})\\
%		\Utility_\_\max\nolimits^y(a) &= -W_{ay} \Delta(\hat{a})\\
%	\forall b \in \ell, \Delta(b) &= W_{ab}\Delta(\hat{a})\\
%	\forall b \in \ell, \Delta(\hat{b}) &=
%	\begin{cases}
%		\frac{\UB(b)}{\UB(b)-\LB(b)}\Delta(b),  &\text{if }W_{bx} > 0\\
%		\max(\Delta(b),-\sol(b)),  &\text{if }  W_{bx} < 0 \text{ and } \sol(b)\geq0\\
%		\max(\Delta(b)+\sol(b),0),  &\text{if }  W_{bx} < 0 \text{ and } \sol(b)<0		 
%	\end{cases}\\
%	\Utility\_\max\nolimits^x(a) &= -\sum_{b \in \ell} W_{bx} \Delta(\hat{b})\\
%	\Utility\_\max\nolimits^z(a) &= \Utility\_\max\nolimits^x(a)+\Utility\_\max\nolimits^y(a)
%\end{align*}
%
%	
%	
%		
%	
%
%
%\section{Formula for Utility}
%
%
%Recall that for nodes $a,b$, we use 
%$\hat{a}, \hat{b}$ to denote variable after ReLU functions, and
%$\Delta(a),\Delta(\hat{a}),\Delta(b),\Delta(\hat{b})$ to denote the changes of those variables. $a$ means the source node, and $b$ means nodes one layer after $a$, and $z$ is 2 layers after $a$ and one after $b$.
%
%The utility function for a node $a$ wrt neuron $z$ is defined inductively as follows:
%
%
%\begin{align*}
%	\Delta(\hat{a}) &= \ReLU(\sol(a))-\sol(\hat{a})\\
%	\Delta(b) &= W_{ab}\Delta(\hat{a})\\
%	\Delta(\hat{b}) &=
%	\begin{cases}
%		\frac{\UB(b)}{\UB(b)-\LB(b)}\Delta(b),  &\text{if }W_{bz} > 0\\
%		\max(\Delta(b),-\sol(b)),  &\text{if }  W_{bz} < 0 \text{ and } \sol(b)\geq0\\
%		\max(\Delta(b)+\sol(b),0),  &\text{if }  W_{bz} < 0 \text{ and } \sol(b)<0		 
%	\end{cases}\\
%	\Utility\_\max\nolimits^z(a) &= -\sum_{b \in \ell} W_{bz} \Delta(\hat{b})
%\end{align*}
%
%}
